A user-extensible and adaptable parser architecture

### 2008-05-25 papers/sgai-2008/paper.tex
Fleshed out the rules section, mainly with content taken from logparser.pdf. Made it clear that rules cause cations to be invoked, based on feedback from     Ariane that it I gave the opposite impression at first. Added a little more to the framework section.

The architecture described herein was developed as part of a larger project to
improve anti-spam techniques by analysing the performance of the set of filters
currently in use, optimising the order and membership of the set based on that
analysis, and developing new measures to supplement the existing filters where
there is a deficiency.  The approach chosen was to analyse the log files
produced by the mail server, rather than modifying the mail server to generate
performance statistics; this approach improves the chances of other sites
testing and utilising the software.  The need arose for a parser capable of
dealing with the huge variety of log lines produced by Postfix~\cite{postfix},
the \MTA{}:

@@

    \item The mail administrator can define custom rejection messages, and
        thus custom log lines.

@@

@@

@@

@@

\cite{postfix}

>>>>

The architecture described herein was developed as part of a larger project to
improve anti-spam techniques by analysing the performance of the set of filters
currently in use, optimising the order and membership of the set based on that
analysis, and developing new measures to supplement the existing filters where
there is a deficiency.  The approach chosen was to analyse the log files
produced by the mail server, rather than modifying the mail server to generate
performance statistics; this approach improves the chances of other sites
testing and utilising the software.  The need arose for a parser capable of
dealing with the huge variety of log lines produced by Postfix~\cite{postfix},
the \MTA{} in use:

@@

    \item The mail administrator can define custom rejection messages,
        resulting in custom log lines.

@@

The framework tries each rule in turn until it finds one which matches the
input line, then invokes the action specified by that rule.

@@

The function tracks the number of times each rule matches; this data is
used during subsequent parsing to optimise the order the rules are tried
in.  Rules specify the Postfix component which produces the log lines they
parse; this avoids needlessly trying rules which won't match the log line,
or worse, might match unintentionally.

@@

Rules are responsible for categorising input lines, specifying both the
data to be extracted from each line and the action that should be invoked.
This implementation stores the rules in the same \SQL{} database that the
results are stored in; this removes any doubt as to which set of rules was
used to produce a set of results.  Other implementations are free to store
the rules in whatever fashion suits their needs.  Each rule defines several
attributes:

    \item [program] The Postfix program whose log lines the rule applies
        to.  This avoids needlessly trying rules which won't match the log
        line, or worse, might match unintentionally.  Rules whose program
        is \texttt{*} will be tried against any log lines which aren't
        parsed by program specific rules.

    \item [regex] The \regex{} to match the log line against.  The \regex{}
        will first have several keywords expanded: this simplifies reading
        and writing rules; avoids needless repetition of complex \regex{}
        components; allows the components to be corrected and/or improved
        in one location; and makes each \regex{} largely self-documenting.

        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}, with the impact of
        compiling and caching \regexes{} covered in \sectionref{Caching
        each regex}.

    \item [result\_cols] Specifies how the fields in the
        log line will be extracted.  The format is:
        \newline \tab{} \texttt{smtp\_code = 1; recipient = 2, sender = 4;}
        \newline i.e.\ semi-colon or comma separated assignment statements,
        with the variable name on the left and the matching capture from
        the \regex{} on the right hand side.

    \item [result\_data] Sometimes rules need to supply a piece of data
        which isn't present in the log line: e.g.\ setting
        \texttt{smtp\_code} when mail is accepted.  The format and allowed
        variables are the same as for \texttt{result\_cols}, except that
        arbitrary data\footnote{Commas and semi-colons cannot be escaped
        and thus cannot be used.  This is intended for use with small
        amounts of data rather than large amounts in any one rule, so
        dealing with escape sequences was deemed unnecessary.} is permitted
        on the right hand side of the assignment.

    \item [action] The action that will be invoked when this rule matches a
        log line; actions are describe in \sectionref{Actions}.

    \item [hits] This counter starts at zero and is incremented each time
        the rule successfully matches.  At the start of each run the parser
        sorts the rules in descending order of hits, and at the end of the
        run updates every rule's hits.  Assuming that the distribution of
        log lines is reasonably consistent between log files, rules
        matching more commonly occurring log lines will be tried before
        rules matching less commonly occurring log lines, lowering the
        program's execution time.  Rule ordering for efficiency is
        discussed in \sectionref{rule ordering for efficiency}.

    \item [priority] This is the user-configurable companion to hits: rules
        will be tried in order of priority, overriding hits.  This allows
        more specific rules to take precedence over more general rules
        (described in \sectionref{overlapping rules}).

The rules used by the Postfix log file parser define several additional
attributes; each implementation is free to add any additional attributes it
requires.

\subsubsection{Overlapping rules}

\label{overlapping rules}

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.\footnote{It may be possible to parse
each rule's regex and determine if any overlap.  The author has not
attempted to do this: given the tremendous power and flexibility of
regex, particularly Perl's (XXX REFERENCE FOR PERL\@; GLOSSARY ENTRY?)
regex, it would be an extremely difficult task.} Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the order in which rules are tried against each line may change between log
files, and the first matching rule wins.  Overlapping rules are frequently
a requirement, allowing a more specific rule to match some log lines and a
more general rule to match the majority, e.g.\ separating \SMTP{} delivery
to specific sites from \SMTP{} delivery to the rest of the world.  The
algorithm provides a facility for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches when parsing the previous log file.  Negative
priorities may be useful for catchall rules.

Detecting overlapping rules is difficult, but the following approaches may
be helpful:

    \item Sort by \regex{} and visually inspect the list, e.g.\ with \SQL{}
        similar to: \textbf{select regex from rules order by regex;}

    \item Compare the results of parsing using sorted, shuffled and
        reversed rules.\footnote{See \sectionref{rule efficiency} for more
        details of sorting the rules.}  Parse a number of log files using
        normal sorting, then dump a textual representation of the rules,
        connections and results tables.  Repeat with shuffled and reversed
        rules, starting with a fresh database.  If there are no overlapping
        rules the tables from each run will be identical; differences
        indicate overlapping rules.  Which rules overlap can be determined
        by examining the differences in the tables: each result contains a
        reference to the rule which created it, if the references differ
        between runs the two rules referenced in the differing records
        overlap.  Unfortunately this method cannot prove the absence of
        overlapping rules; it can detect overlapping rules, but only if
        there are log lines in the input files which match more than one
        rule.

@@
